---
title: "P9120 HW4"
output: html_document
date: '2022-12-07'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(Matrix)
library(glmnet)
library(keras)
library(jpeg)
library(tidyverse)
library(ggthemes)
```

## Exercise 8

From your collection of personal photographs, pick 10 images of animals (such as dogs, cats, birds, farm animals, etc.). If the subject does not occupy a reasonable part of the image, then crop the image. Now use a pretrained image classification CNN as in Lab 10.9.4 to predict the class of each of your images, and report the probabilities for the top five predicted classes for each image.

Construct the Pretrained CNN model:

```{r}
img_dir <- "book_images"
image_names <- list.files(img_dir)
image_names
num_images <- length(image_names)
x <- array(dim = c(num_images, 224, 224, 3)) 

for (i in 1:num_images) {
  img_path <- paste(img_dir, image_names[i], sep = "/") 
  img <- image_load(img_path, target_size = c(224, 224))
  x[i,,, ] <- image_to_array(img)
}

```

Look at the 10 images:

```{r}
par(mar = c(0, 0, 0, 0), mfrow = c(2, 5))
x_temp <- x / 255
for (i in 1:10) plot(as.raster(x_temp[i,,, ]))
```

Preprocess the image:

```{r}
x <- imagenet_preprocess_input(x)
```

Load the Pretrained Model：

```{r}
model <- application_resnet50(weights = "imagenet") 
summary(model)
```

Print the top 5 prediction

```{r}
pred10 <- model %>% predict(x) %>%
  imagenet_decode_predictions(top = 5) 
names(pred10) <- image_names
print(pred10)
```

The prediction is generally correct. From the above we know that the background of the image may have an impact on the prediction. e.g. the last image is actually a white swan in a dark background. 

## Exercise 13

On the book website, www.statlearning.com, there is a gene expression data set (Ch12Ex13.csv) that consists of 40 tissue samples with measurements on 1,000 genes. The first 20 samples are from healthy patients, while the second 20 are from a diseased group.

(a)  Load in the data using `read.csv()`. You will need to select `header = F`.

```{r }
set.seed(1)
df <- read.csv("Ch12Ex13.csv", header = FALSE)
head(df)
```
The 40 columns represent 40 patients, and the rows are different genes.

(b) Apply hierarchical clustering to the samples using correlation-based distance, and plot the dendrogram. Do the genes separate the samples into the two groups? Do your results depend on the type of linkage used?

We are using correlation-based distance as the dissimilarity measure
```{r}
set.seed(2022)
corrplot::corrplot(cor(df), method = 'color', 
         order = 'hclust', hclust.method = 'complete', 
         tl.col = 'black', tl.cex = 0.7)
dists <- as.dist(1 - cor(df))
```

Try with different linkages:
```{r}
methods <- c('centroid', 'average', 'single', 'complete')
set.seed(2022)
par(mfrow = c(2,2))
for (method in methods) {
    clusts <- hclust(dists, method = method)
    
    plot(clusts, 
         col = "light blue",
         lwd = 3, lty = 1, 
         sub = "", hang = -1, 
         axes = FALSE,
         main = paste0('Cluster Dendrogram using ', method, ' linkage'))
}
```

All the clustering results are different. The centroid, complete and single linkage methods all separate the patients into 2 main clusters. However, the average linkage method sepearate the patients into 3 main clusters.

We see that the first 20 patients are clustered into 1 group and the last 20 are clustered into the second group, consistent with our knowledge.

(c) Your collaborator wants to know which genes differ the most across the two groups. Suggest a way to answer this question, and apply it here.

I believe we can use PCA to find the genes differ the most across the two patients groups:

1. Perform PCA on the dataset

2. For the first two principle components, perform K-Means clustering and make plots to see if there is good separation on those components.

3. If there is such separation we will test to see if one group of genes is the one we are looking for.

4. Similarly, perform PCA on the transposed dataset on genes (genes as columns)

5. Check if the samples are well-separated into two main groups for the first two components using k-means

6. If they are well-separated, the genes in the well separated components are most likely correlate with the genes of interest

7. We find the genes in common between step 3 and 6. Those genes are likely to be the most different genes between the two groups.

### Patients as variables

```{r}
set.seed(2022)
require(ggbiplot); require(ggthemes)
# Perform PCA
df.pca <- prcomp(df)
summary(df.pca)
pca_2 <- df.pca$x %>%
    as_tibble() %>%
    dplyr::select(PC1, PC2)
# Perform k-means
pca_kmeans <- pca_2 %>%
    kmeans(centers = 2)
ggbiplot::ggbiplot(df.pca, groups = factor(pca_kmeans$cluster), 
         ellipse = TRUE) +
    geom_point(aes(col = factor(pca_kmeans$cluster)), 
               size = 2, alpha = 0.2) +
    theme(legend.position = 'top') +
    scale_color_manual(name = 'K-Means Group of Patient',
                       values = c('#a6cee3', '#e31a1c')) +
    ggtitle('K-Means Clustering of First Two Principal Components')
```

Checking the result with k-means, we know that the patients are clearly separated into two groups by the first component.

Now we can see which variables correlate most strongly with the first factor:

```{r}
strong_genes1 <- df %>%
    dplyr::mutate(Variable = paste0('Gene ', row_number())) %>%
    filter(pca_kmeans$cluster == 2) %>%
    select(Variable)

strong_genes1 %>% knitr::kable()
```

There are `r nrow(strong_genes1)` genes that are important. Now we look at the PCA with the genes:

```{r}
# tranpose the dataframe
gene_df <- df %>%
    t %>%
    as.tibble
names(gene_df) <- paste0('Gene', 1:1000)
```

### Genes as variables

Perform the PCA and plot the k-means plot:

```{r}
set.seed(2022)

df.pca <- prcomp(gene_df)

gene_pca_2 <- df.pca$x %>%
    as.tibble %>%
    select(PC1, PC2)

gene_pca_kmeans <- gene_pca_2 %>%
    kmeans(centers = 2)

gene_df <- gene_df %>%
    mutate(Patient = paste0('V',1:40))

ggbiplot(df.pca, groups = factor(gene_pca_kmeans$cluster), 
         ellipse = TRUE, var.axes = FALSE, labels = gene_df$Patient,
         labels.size = 5) +
    geom_point(aes(col = factor(gene_pca_kmeans$cluster)), 
               size = 2, alpha = 0.2) +
    theme(legend.position = 'top') +
    scale_color_manual(name = 'K-Means Group of Patient',
                       values = c('#a6cee3', '#e31a1c')) +
    ggtitle('K-Means Clustering of First Two Principal Components')
```

Checking the result with k-means, we know that the patients are clearly separated into two groups by the first component.
They are likely to represent the diseased and non-diseased populations.

We then select the genes in common

```{r}
strong_genes2 <- df.pca$rotation %>%
    as.tibble %>%
    select(PC1) %>%
    mutate(Variable = paste0('Gene ', 1:1000)) %>%
    arrange(desc(abs(PC1))) %>%
    select(Variable) %>%
    head(110)
```


```{r}
common_genes <- strong_genes1 %>%
    inner_join(strong_genes2)

nrow(common_genes)
```

They perfectly overlapped. The 110 genes are likely to be the ones we are looking for. Arranged descending by the difference, they are:

```{r}
strong_genes2 %>% knitr::kable()
```




